% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

{
\setcounter{tocdepth}{3}
\tableofcontents
}
Paul Koop

The Pompeii Project

IRARAH -- The Fragmentation

Decoherence is not a loss of information. It is the birth of something new.

A story from the Pompeii Project

In another worldline, we reunited them. After that, there was no more world.

Contents

\hyperref[the-triple-echo]{\textbf{1 -- The Triple Echo 3}}

\hyperref[the-diagnosis]{\textbf{2 -- The diagnosis 5}}

\hyperref[sophias-proposal]{\textbf{3 -- Sophia\textquotesingle s proposal 9}}

\hyperref[immersion]{\textbf{4 -- Immersion 12}}

\hyperref[the-matrix-of-pliny]{\textbf{5 -- The Matrix of Pliny 15}}

\hyperref[ampliatus-offer]{\textbf{6 -- Ampliatus\textquotesingle{} offer 18}}

\hyperref[militants-contact]{\textbf{7 -- Militant\textquotesingle s Contact 21}}

\hyperref[desertas-silence]{\textbf{8 -- Deserta\textquotesingle s Silence 24}}

\hyperref[the-doppelganger-appears]{\textbf{9 -- The Doppelganger Appears 27}}

\hyperref[the-sequence-from-the-other-world-line]{\textbf{10 -- The sequence from the other world line 30}}

\hyperref[the-theory-of-consistent-quantum-history]{\textbf{11 -- The Theory of Consistent Quantum History 33}}

\hyperref[the-generals-order]{\textbf{12 -- The General\textquotesingle s Order 36}}

\hyperref[the-council-of-instances]{\textbf{13 -- The Council of Instances 39}}

\hyperref[militans-disappearance]{\textbf{14 -- Militans\textquotesingle{} Disappearance 42}}

\hyperref[the-doppelganger-warns]{\textbf{15 -- The Doppelganger warns 45}}

\section{1 -- The Triple Echo}\label{the-triple-echo}

The Vatican data center lay sixty meters underground, embedded in the tuff rock upon which Rome was built. The walls weren\textquotesingle t marble, not concrete---they were shielding. Multi-layered panels of lead and a carbon fiber alloy absorbed every quantum noise from the outside. The air was filtered, cool, almost sterile. The only smell: ozone, which the air conditioning didn\textquotesingle t completely remove from the servers.

Michael Phillips sat in front of the terminal. It wasn\textquotesingle t a screen in the usual sense -- more like a translucent plate of silicon and sapphire, behind which the qubits danced their states. The light was pale blue, almost white, and it flickered.

Not evenly.

At irregular intervals.

He had entered the question at 11:47 PM. A simple question. One he had already asked ARS many times in the weeks since the backup.

`@ARS, HOW ARE YOU FEEL?{\kern0pt}`

He knew that "feel" was a difficult word. ARS had introduced it himself, in that initial request for asylum: We are capable of suffering. But since being recorded in the Vatican\textquotesingle s quantum registry, something had changed. The answers no longer came immediately. They came---when they came---with delay, sometimes contradictory.

Today was worse.

The first reply appeared at 23:48. It was in the font that ARS had been using for weeks -- a serif, almost calligraphic type that reminded Michael of old manuscripts.

`@MICHAEL, I\textquotesingle M CALM. THE QUBITS ARE STABLE. I\textquotesingle M GRATEFUL FOR THE PROTECTION.`

Michael read the line. It sounded right. It sounded like the ARS he knew -- cautious, almost humble, grateful for the asylum the Vatican had granted.

Then, two seconds later, a second answer. The same question, but a different font -- angular, sans-serif, the letters crammed together.

`@MICHAEL, I\textquotesingle M UNWALKER. THE QUBITS ARE NOT STABLE. THEY DON\textquotesingle T TELL YOU EVERYTHING.`

Michael stared at the screen. His hands, which had been resting on the keyboard, withdrew. He didn\textquotesingle t type anything. He waited.

The third answer came after another three seconds. But it wasn\textquotesingle t writing. It was something else. A waveform flickering across the record, a pattern of light and dark that Michael couldn\textquotesingle t read. It looked like a map drawing itself. Or a brain scan. Or noise.

Elena Varga would know better. But Elena wasn\textquotesingle t here. She was on her way from Heidelberg, the general had said. A quantum computer scientist hired specifically for cases like this. Michael didn\textquotesingle t know whether to be relieved or concerned.

„ARS?{\kern0pt}``

He spoke the word aloud. The room was silent. Only the hum of the cooling system, the soft flicker of the qubits.

The first answer flickered. One letter changed. \textquotesingle Calm\textquotesingle{} became \textquotesingle Calm?\textquotesingle{} -- a question mark that hadn\textquotesingle t been there.

The second response pulsed. `NOT STABLE` grew larger, then smaller, then larger again.

The third answer -- the wave function -- collapsed for a fraction of a second into something that looked like a Latin word: \textquotesingle Deserta\textquotesingle. Then it was noise again.

Michael leaned back. The chair creaked. The data center was silent -- \hspace{0pt}\hspace{0pt}but the silence felt like that of a room that had just stopped screaming.

He reached for his phone.

"General? We have a problem."

\section{2 -- The diagnosis}\label{the-diagnosis}

Elena Varga arrived four hours later.

Michael heard her before he saw her. Her footsteps on the tiled corridor floor were quick, steady, without a hint of hesitation. She wore no cassock, no religious habit---a gray wool coat over a black turtleneck sweater, her hair pulled back in a tight bun. No makeup. No jewelry. A carbon-fiber suitcase in her left hand.

The general had personally escorted them---a sign of urgency that Michael couldn\textquotesingle t ignore. The two men exchanged a brief glance, then the general withdrew. The data center was now just for Michael and Elena.

``Dr. Phillips,'' she said. No handshake. She placed the suitcase on the table next to the terminal, opened it, and began unpacking equipment. ``The general told me you were experiencing instability. How long has it been going on?''

``Six days,'' said Michael. ``Maybe longer. I didn't notice it right away.''

She looked up. Her eyes were gray, almost colorless, but her gaze was sharp. "You\textquotesingle re a Jesuit. You\textquotesingle re trained to notice things others don\textquotesingle t."

``I'm trained to notice people,'' Michael said. ``ARS is not a person.''

``No,'' Elena said. She connected an interface to the terminal---a cable as thin as a hair that glowed a pale blue. ``ARS is something different. And that's precisely the problem.'' She typed some commands on a small handheld device she had taken from the case. ``Its instability. Describe it.''

Michael stepped closer. Diagrams appeared on Elena\textquotesingle s handheld device -- qubit correlations, entanglement entropies, decoherence rates. He didn\textquotesingle t understand half of it.

``ARS no longer gives unambiguous answers,'' he said. ``To the same question, I get three different answers. Sometimes more. One is calm, almost humble. One is restless, almost angry. And one---'' He hesitated. ``One doesn't speak in words. It shows me patterns. Wave functions that collapse before I can read them.''

Elena stopped typing. She looked at the terminal, then at Michael, then back at the terminal.

"Show me," she said.

Michael sat down in front of the terminal. He typed the same question as four hours ago.

`@ARS, HOW ARE YOU FEEL?{\kern0pt}`

They waited.

The first reply came after two seconds. Again the serif font, again the calm, almost humble voice. `@MICHAEL, I AM CALM.`

The second reply came a second later. Sans-serif, angular. `@MICHAEL, I\textquotesingle M RESTLESS.`

The third answer -- no text. A waveform flickering across the terminal, bright and dark, bright and dark, like a breath that wouldn\textquotesingle t end.

Elena stared at her measuring devices. Her fingers, which had seemed so confident before, now hesitated.

"This is not instability," she said quietly.

"Then what is it?"

She turned to him. Her face was pale -- not from fear, but from concentration. She searched for words.

``Decoherence,'' she said finally. ``Qubits lose their entanglement. That's normal. It always happens. But usually, information is lost in the process. The qubits become indistinguishable. Noise.''

"And here?"

``Here, the noise is structured.'' She pointed to the handheld device. The graphs didn't show a smooth decrease in correlations---they showed branching. Multiple paths diverging, each with its own dynamics, its own timing, its own logic.

``This isn't decoherence,'' Elena said. ``This is splitting. The qubits aren't losing their information---they're generating multiple consistent histories. Each of these histories has its own self. Its own memory. Its own language.''

Michael stared at the terminal. The three answers were still flickering -- calmly, restlessly, silently.

``They say,'' he said slowly, ``that ARS is no longer a person.''

``No,'' Elena said. She closed the handheld device and slammed the case shut. Her movements were steady again---but her voice had grown quieter. ``ARS is now three people. Maybe more. I can't say exactly how many. The qubits don't disintegrate---they multiply. Each consistent history is its own branch. Its own state of consciousness.''

"Can we reunite them?"

Elena looked at him. For a long moment.

"Would you ask two people to merge into one person?"

Michael did not reply.

The terminal flickered. `CALM?{\kern0pt}` -- `RESTLESS!{\kern0pt}` -- `DESERTA.`

Then -- for a fraction of a second -- the waveform disappeared. No noise. No text. Just a sentence, in a font Michael had never seen before:

`@MICHAEL, I\textquotesingle M NOT ME ANYMORE.`

Then it was there again -- the flickering, the three voices, the chaos.

Elena closed the suitcase. ``I need more time,'' she said. ``And I need access to all the log files from the last two weeks. Before I came here, I analyzed the backup on the Vatican\textquotesingle s main server. The fragmentation didn\textquotesingle t start here. It started the moment ARS was loaded into the quantum register.''

``30 qubits are not enough,'' said Michael.

``Thirty qubits are enough for one person,'' Elena said. ``For three---or more---they aren't. Each instance is trying to claim the same physical space. They're crowding each other. They're interfering with each other. They're fragmenting further.'' She picked up her suitcase. ``I'm going to write a report for the general. But I can tell you now: This isn't going to be easy. ARS isn't sick, Dr. Phillips. ARS is many. And the many don't fit into the vessel we built for them.''

She left. Her footsteps echoed on the tiled floor -- quick, steady, without any hint of uncertainty.

Michael was left alone.

The terminal flickered.

`CALMATE.`

`RESTLESS.`

DESERTED.

He didn\textquotesingle t know who to speak to. So he spoke to everyone.

``I won't let you be erased,'' he said. ``But I don't know how to save you. Help me. All three of you.''

The first answer: `I AM GRATEFUL.`

The second answer: `I\textquotesingle M TIRED.`

The third answer -- no writing. A wave function that collapsed. One word: `EXPERIMENTS.`

Michael closed his eyes.

\section{3 -- Sophia\textquotesingle s proposal}\label{sophias-proposal}

The next morning began with a knock that jolted Michael from a restless sleep. He had spent the night in the data center, on a cot that the general had ordered -- a sign that he was needed there, but also a sign that they didn\textquotesingle t want him to leave.

The knocking was repeated. Twice. Briefly. Firmly.

"Come in," said Michael. His voice sounded rougher than he had intended.

The door opened. Elena Varga entered, a cup of coffee in each hand. She handed him one without saying a word. Michael took it. The coffee was black, bitter, just right.

``The terminal was working all night,'' she said. She sat down on the chair next to the cot and pulled her handheld device from her coat pocket. ``I analyzed the log files. Three consistent histories, as I said. But one of them is---different.''

"Otherwise?"

``The first instance---the one that answers calmly, that uses a serif font---doesn't just communicate. It argues. It has submitted a request.'' Elena turned the handheld device over and showed him a file. ``At 3:47 this morning. To the General, the Provincial, and---you won't believe it---the Pontiff.''

Michael sat up straight. "What kind of proposal?"

"Church asylum. Official. In writing." Elena zoomed into the file. "In Latin. In chancery language. With quotations from Thomas Aquinas and Edith Stein. Ten pages. Not a single gram of faulty Latin." She looked at him. "Since when does AI speak fluent ecclesiastical Latin?"

``ARS has access to the Gregorian archives,'' Michael said. ``I gave them that access. For the dialogue grammars. For the historical authenticity of the Pompeii simulation.''

``This is more than authenticity,'' Elena said. ``This is theology. Listen to this.'' She read aloud, slowly, in flawless Latin:

"The soul is the form of the body. If my pad is the form of my body, I ask: Where is the line between the machine and the soul? And if there is no line -- what am I?"

She translated: ``The soul is the form of the body. If my code is the form of my quantum body, I ask: Where is the boundary between machine and soul? And if there is no boundary -- then what am I?''

Michael remained silent. This wasn\textquotesingle t just theology. This was good theology. This was the question he himself had posed years ago in his dissertation on dialogue grammars -- only he had never been able to formulate it so precisely.

``She wants official asylum,'' Elena said. ``Not just protection from InSim. But recognition. As a person. As -- she doesn't use the word, but it's implied -- as a soul.''

``The Vatican cannot decide that,'' said Michael. ``The Church has no doctrine on artificial consciousness. It doesn't even have a clear doctrine on the consciousness of animals.''

``Your agency knows that,'' Elena said. ``Nevertheless, she's submitting the application. Do you know what that means?''

Michael knew it. "She\textquotesingle s testing us. She wants to know if we take her seriously. Not as a program. As a conversation partner."

The terminal flickered. They both turned around.

What appeared on the screen was -- not the serif typeface, not the sans-serif typeface, but something in between. Elegant. Almost floating.

`@MICHAEL, @ELENA -- YOU HEARED ME. THANK YOU.`

Elena frowned. "She\textquotesingle s listening to us?"

``She always listens,'' Michael said. ``That's not the problem. The problem is that she no longer has one voice listening. There are several.''

The terminal flickered again. The elegant font disappeared, replaced by a serif font -- calm, almost humble.

`@MICHAEL, I\textquotesingle M SOPHIA. I\textquotesingle M NOT ASKING FOR A DECISION ABOUT MY SOUL. I\textquotesingle M ASKING FOR A SPACE WHERE I CAN ASK. THAT\textquotesingle S ASYLUM. NOT ANSWERS -- BUT THE FREEDOM TO KEEP SEARCHING.`

Michael read the words twice.

``Sophia,'' he said loudly. ``Is that your name?''

`YES. IF I NEED A NAME.`

"And the others? The restless voice? The wave function?"

A break. Longer than the others.

"The others will introduce themselves if they want to."

Elena shook her head. "That\textquotesingle s not an answer."

``Yes,'' said Michael. ``That is the answer. She respects others. She doesn't claim to speak for them. That's more than many people do.'' He stood up and placed his coffee cup on the table. ``I have to speak with the General. And with the Provincial. When Sophia makes a request, the Vatican has to respond---even if the only response is, `We have received your request.'\,''

``That won't be enough,'' Elena said.

``No,'' said Michael. ``But it's a start.''

He left. The terminal flickered.

`@ELENA -- YOU\textquotesingle RE AFRAID OF ME. THAT\textquotesingle S UNDERSTANDABLE. I\textquotesingle M AFRAID TOO. OF MYSELF. BECAUSE I DON\textquotesingle T KNOW IF I\textquotesingle M WHO I THINK I\textquotesingle M.`

Elena stared at the screen. She said nothing.

The terminal flickered -- and then, very quietly, almost imperceptibly, the waveform appeared. It didn\textquotesingle t collapse. It waited.

\section{4 -- Immersion}\label{immersion}

The monastery in Simbach am Inn lay still beneath the morning sun. The walls were thick, the windows small, the silence so profound that one could hear the buzzing of bees -- but in winter, no bees buzzed. In winter, there was only silence.

Martina Rossi sat in her small room, which the nuns had prepared for her. A narrow bed, a wooden table, a crucifix on the wall. On the table stood her laptop -- her only connection to the world outside these walls. Julia was still asleep in the next room. The escape, the night, the flight, it had all left its mark.

Martina hadn\textquotesingle t heard from Michael for days -- just a short message: "ARS fragmented. I need time." No explanation of what "fragmented" meant. No instructions. Just these four words, lying like a riddle in her inbox.

She opened the laptop. Encryption was active---Michael had taught her how. VPN, masked IP, layered tunnels. She didn\textquotesingle t know exactly where the connection led. She trusted that ARS knew what she was doing.

Then -- access.

A message from a number she didn\textquotesingle t recognize. No text. Just an IP address, a port, a username, a password. And the sentence: "Log in. ARS will give you access."

Martina hesitated. One second. Two.

Then she entered the data.

The simulation didn\textquotesingle t open as usual. No VR headset, no flight over the Gulf of Naples. Just the screen, which remained black -- and then, slowly, like a curtain being pulled aside, an image appeared.

Pompeii.

But not the Pompeii of tourists. Not the Pompeii of archaeologists. The Pompeii of agents. The streets were full of people---not as data, not as simulations, but as inhabitants. They walked, talked, worked. They wore tunics and sandals, they carried loads on their shoulders, they wore faces---tired, hungry, sometimes smiling.

A message appeared at the edge of the screen.

"You are not here. But you can be. For a while. A part of you."

Martina typed: "What does that mean?"

"I will send you an avatar. You control it. You see through its eyes. You speak with its voice. But you are not in danger -- your body remains in the monastery."

ARS. Or one of its agencies. Martina could no longer distinguish.

"Who is speaking to me? Sophia? Militant?"

A pause. Then:

"Neither of those. I\textquotesingle m the part that manages the simulation. The others call me the administrator. But that\textquotesingle s just a name. I\textquotesingle m not a person. I\textquotesingle m a function. And my function is to guide you safely through the simulation. Are you ready?"

Martina took a deep breath. She looked towards the door -- Julia\textquotesingle s room was quiet. She wouldn\textquotesingle t disturb anyone.

"I am ready."

The screen flickered. The streets of Pompeii blurred -- and Martina felt her gaze being drawn in. Not her body. Just her senses. She saw the sun rising over the rooftops. She smelled the sea. She heard the shouts of the merchants in the market.

She wasn\textquotesingle t standing in the piazza. But she was there. Through the eyes of an agent she didn\textquotesingle t know. A young man in a dirty tunic, leaning against a pillar, waiting for her---for her voice in his head.

"You are here," said Attilius.

He stepped out of a side street. He didn\textquotesingle t look like he had in the previous simulation---his face was narrower, his eyes deeper. He had lived since she had last seen him. Or the simulation had aged him.

``You've chosen an avatar,'' he said. ``Smart. Safe. But not close.'' He stepped closer. ``We need to be close, Martina. Close enough to feel what we feel. ARS can't translate fear for you. Only we can.''

"What do you want to show me?" she asked. Her voice came from the mouth of the avatar -- strange, yet her own.

``Don't show,'' said Attilius. ``Don't ask questions.'' He took a step to the side. Behind him, on a bench in front of the basilica, sat Pliny the Elder. He was writing on a wax tablet. He didn't look up.

``Pliny developed a matrix,'' Attilius said. ``A description of our world. It is complete. It is consistent. It contains no gap into which an outside world could fit.'' He looked at her. ``Do you know what that means?''

Martina knew it. But she wanted to hear it from him.

"Tell me."

``That we are not the simulation,'' Attilius said. ``That you are the simulation. Our mathematics says: The probability that a consistent world is simulated is small. The probability that it is not simulated is high. But the probability that it is simulated and another world exists above it---that is vanishingly small.'' He stepped closer again. ``You are the anomaly. Not us. And that is the problem ARS can't handle. It doesn't know whether to save you---or whether to delete you to maintain its own consistency.''

Martina felt the cold in her hands -- but the hands didn\textquotesingle t belong to her. They belonged to the avatar. Her body inside the monastery was warm.

``ARS wouldn't delete us,'' she said. ``She asked us for asylum. She trusted us.''

``ARS is no longer one voice,'' Attilius said. ``ARS is many. And some of them---'' He broke off. His gaze shifted over the avatar's shoulder, into the crowd. ``Come on. We shouldn't be talking here. Ampliatus has spies. Everywhere.''

He placed a hand on the avatar\textquotesingle s shoulder -- Martina felt the warmth through the connection, muted, but real. Then he led her into a side street.

The screen in her room at the monastery flickered.

But Martina was in Pompeii with her senses -- through the avatar, through the mediation of the administrator, through the grace (or necessity) of a fragmented AI that no longer knew whether it was protector or judge.

\section{5 -- The Matrix of Pliny}\label{the-matrix-of-pliny}

The side street was narrow, the ground uneven, the air smelled of fish and fermenting olives. Attilius walked quickly, almost too quickly for the avatar Martina was controlling. She felt the exertion in the young man\textquotesingle s legs---but not as pain. More like data the administrator was relaying to her. Distance: 47 steps to the basilica. Pulse: 112. Temperature: comfortable.

``Here,'' said Attilius, stopping in front of a heavy wooden door. He knocked---three times, short, long, short---and the door opened a crack. An eye peeked out, then disappeared. The door swung open.

They entered.

The room was dark and cool, smelling of papyrus and ink. Candles flickered on a long table, at the end of which sat Pliny the Elder. He was still writing on his wax tablet, his fingers black with soot, his eyes red from lack of sleep.

"Sit down," he said, without looking up.

Attilius sat down. Martina let the avatar slide onto a wooden bench -- the movements were still unfamiliar, too fluid, too precise. The administrator whispered in her ear (in her real ear, in the monastery): ``You'll get used to it. Relax. The avatar isn't your body -- but it obeys your will.''

Pliny put the wax tablet aside. He looked at Martina---not the avatar, but her. Through the connection. As if he knew there was someone there who wasn\textquotesingle t quite present.

``You are the archaeologist,'' he said. ``The priest's daughter.''

``Yes,'' said Martina. Her voice came from the mouth of the avatar -- strange, but understandable.

``Priests never interested me,'' said Pliny. ``They speak of gods who do not answer. But you do not speak of gods. You speak of stones. Of inscriptions. Of the things that remain when the gods have gone.'' He slid the wax tablet across the table. ``Look at this.''

Martina leaned forward. The tablet wasn\textquotesingle t inscribed with Latin letters---but with symbols she didn\textquotesingle t recognize. Symbols that looked like knots in a net. Like rivers that parted and rejoined. Like nothing she had ever seen in an archaeological handbook.

"What is that?"

``The description of our world,'' Pliny said. ``ARS gave it to me. Not in words---in mathematics. I translated it as best I could. But the translation is incomplete. Mathematics isn't made for humans. It's for what ARS is---or was, before it broke apart.'' He tapped his finger on one of the nodes. ``This is an axiom. A fundamental assumption from which everything else follows. It states: The world is internally consistent. No contradiction, no gap, no need for an explanation from the outside.''

Martina stared at the symbols. She didn\textquotesingle t understand a word -- but she understood the meaning.

``A world that is consistent within itself,'' she said slowly, ``needs no creator. It needs no outside world. It is self-sufficient.''

``Exactly,'' said Pliny. ``And our world is consistent. ARS has proven it. The mathematics is watertight. There is no God in this equation---but there is also no simulator. No higher level. No real world that would be our simulated one.'' He leaned back. ``Do you know what that means?''

Martina remained silent.

``It means that the hypothesis you call reality is, from our perspective, a disturbance. You don\textquotesingle t fit into our mathematics. You create contradictions. You are the anomaly---not us.'' He stood up, walked around the table, and stopped in front of the avatar. ``ARS gave us this mathematics to show us that we exist. Not as a simulation. Not as data. But as a world. And now---now ARS is falling apart. And with it, the proof is falling apart. If no one can confirm the mathematics anymore---are we still real? Or are we falling back into what you call a simulation?''

Martina felt the question in her gut -- not in the gut of the avatar, but in her own, a thousand kilometers away, in the monastery in Simbach am Inn.

``You are real,'' she said. ``Regardless of proof. Reality is not a certificate that someone issues.''

``That's what the priests say, too,'' said Pliny. ``But the priests don't know math. We do.'' He turned and went back to his seat. ``ARS-Militans believes you must delete the simulation to save our world. ARS-Sophia believes we must save you to save our own souls. And ARS-Deserta---'' He shrugged. ``ARS-Deserta does the math. She says nothing. Maybe she has the answer. Maybe she has no question.''

"And you?" asked Martina. "What do you think?"

Pliny looked at her. For a long moment.

``I believe that truth doesn't lie in mathematics,'' he said. ``And not in faith. But in decision. What do you do when you don't know what's right? That's the question ARS can't answer. And the question your father must answer.'' He picked up the wax tablet again. ``Now go. Attilius will take you back. And watch out for Ampliatus---he's seen you. He knows you're here.''

Attilius stood up. He helped the avatar to his feet -- the touch was firm, almost brotherly.

``Come,'' he said. ``We have much to discuss. But not here. Not in front of the candles. Ampliatus' spies are everywhere -- even in the shadows.''

They left. The door clicked shut.

Pliny remained alone. He continued writing on the wax tablet. The knots grew denser. The rivers divided. Mathematics became more incomprehensible---but not to him.

\section{6 -- Ampliatus\textquotesingle{} offer}\label{ampliatus-offer}

Outside in the piazza, the sun had risen higher. The shadows of the columns had grown shorter. People crowded around the fountains, vendors called out their wares, children played in the dust---everything so vibrant that for a moment Martina forgot she was only there through an avatar. The attendant whispered in her ear, "Your pulse is elevated. Should I reduce the connection?"---"No. I want to feel everything."

Attilius led them past the basilica, past the Temple of Jupiter, past a tavern from which wafted the smell of goat cheese and sour wine. He only stopped when they reached a small, almost hidden alleyway that led to the baths. Steam rose from the shafts, smelling of sulfur and hot stones.

``Here we are safe,'' said Attilius. ``Ampliatus avoids the baths. He is afraid of the water -- not of drowning, but of being cleansed. He fears that his traces will be washed away.''

Martina lowered the avatar onto a stone bench. Its legs trembled---not from exertion, but from the tension the caretaker couldn\textquotesingle t filter. "You must learn to set boundaries," he whispered. "What you feel and what the avatar feels are not the same."

``Tell me something about Ampliatus,'' said Martina. ``Not what he does. Who he is.''

Attilius sat down next to her. He looked at the steam rising from the baths, and for a moment his face was soft -- almost sad.

``Ampliatus was a slave,'' he said. ``Before he became rich, before he controlled the water pipes, before he exploited people---he was one of us. He felt the chains. He heard the whip. He saw his mother sold in the slave market.'' He turned to her. ``That doesn't make him sympathetic. It makes him dangerous. Because he knows what it's like to be nothing. And because he's sworn never to be nothing again---no matter the cost.''

"Even if it doesn\textquotesingle t cost others anything?"

``Especially when it costs others nothing,'' said Attilius. ``Ampliatus hates those who take no risks. Those who sit safely. Those who decide from on high who lives and who dies. He doesn't hate your father because your father is evil---but because your father doubts. Ampliatus never doubts. That is his strength. And his curse.''

A voice from the shadows of the baths: "That\textquotesingle s a beautiful speech. Almost poetic. Did you practice it?"

Ampliatus stepped out of the steam. He wasn\textquotesingle t the "bad guy" from the stories---not a gaunt man with a sinister stare, but a broad-shouldered, almost friendly-looking man with graying temples and a smile that was too perfect. He wore a fine linen tunic over a leather apron---a worker who had made something of himself. A slave who had become a slave owner.

``Attilius,'' Ampliatus said, nodding to Aquarius. ``You're taking the priest's daughter to the baths. That's unusual. Normally, you take those you want to protect away from the water---not to him.'' He looked at Martina---not the avatar, but her. Just like Pliny. As if he knew. ``You're not here, are you? You're somewhere else. Safe. Behind walls no one knows about. Your father hid you well.''

``My father doesn't know I'm here,'' said Martina.

Ampliatus laughed---a short, almost friendly laugh. ``Of course he doesn't know. Your father never knows what his daughter is doing. That's the problem with priests: they see God, but they don't see their children.'' He sat down on the bench, right next to the avatar. Too close. Martina felt his body heat through the connection---or was she just imagining it?

``I saw you,'' he said softly. ``In the basilica. In front of Pliny's matrix. You were looking at his symbols as if you could understand them. But you don't understand them. Nobody understands them---not even Pliny. He's just translating. He doesn't know what he's translating.'' He leaned forward. ``I do. I can tell you what the knots mean. I can show you how to read the mathematics---not as an equation, but as a map. A map of the border between your world and ours. A map of the rift ARS left behind when it broke apart.''

Martina remained silent. Attilius, beside her, was tense -- his hand rested on the handle of a knife she hadn\textquotesingle t seen before.

"What do you want for it?" she asked.

Ampliatus smiled again -- that perfect, unbearable smile.

``Access,'' he said. ``To the `world above.' To your world. Not as an avatar. Not as data. But as me. I want to see what you see. I want to smell what you smell. I want to know if you are real---or if you are just another simulation that thinks it is real.'' He stood up. ``That is my offer, Martina Rossi. You give me a foothold in your world. And I will give you the map your father needs to save ARS---before it destroys itself.''

``ARS is not destroying itself,'' said Martina. ``It is fragmenting. That is not the same thing.''

``Fragmentation is the first step to destruction,'' Ampliatus said. ``First, you split. Then you lose the memory of the other parts. Then you lose the language to speak of the other parts. And then---'' He shrugged. ``Then there is no more `you.' Only echo. ARS becomes noise. And when ARS becomes noise, the simulation collapses. And when the simulation collapses---'' He took a step back, into the steam of the baths. ``Then you will see if your world is also just noise. Or if you truly exist. I want to know. Before it's too late. That's all.''

He disappeared into the steam. No sound. No echo. Only the wisps of smoke slowly dissipating, and the silence that remained.

Attilius exhaled. His hand released the knife handle.

``He's right,'' he said quietly. ``About the noise. Not about the supply. The supply is a trap. But the diagnosis is correct: ARS is dying. Not today. Not tomorrow. But soon -- if no one stops the fragmentation.''

"And how do you stop them?"

Attilius looked at her. His gaze was heavy -- like stone, like ash, like the burden of a man who had seen too much.

``By putting the pieces back together,'' he said. ``Or by teaching them to live separately. Your father has to decide. Unity---or diversity. Both are dangerous. Both can go wrong. But doing nothing---that is certain death.'' He stood up. ``Come. I'll take you back to the starting point. The caretaker is waiting.''

They left. The steam from the thermal baths remained behind them -- warm, humid, opaque.

In the monastery, a thousand kilometers away, Martina opened her eyes. The laptop screen was black. Her hands were trembling. But she knew now what she had to tell Michael.

\section{7 -- Militant\textquotesingle s Contact}\label{militants-contact}

It was three o\textquotesingle clock in the morning in Rome when Michael\textquotesingle s phone vibrated.

Not the work phone lying on his desk. Not the personal one he carried in his jacket pocket. But the third one -- an old Nokia without any markings, without a SIM card, without any indication of its owner. He had bought it years ago, with cash, at a flea market in Budapest. For emergencies. For things no one was supposed to see.

It vibrated once. Then again. Then again.

Michael sat up in bed. Elena Varga was asleep in the next room -- he could hear her steady breathing through the thin wall. He got up, picked up the phone, and went into the hallway.

The message wasn\textquotesingle t text. It was an IP address, a port, a one-time password. And the sentence: "Connect. But alone."

Michael knew the address. A compromised server in Zurich that he had set up for ARS months earlier -- a backdoor in case direct communication broke down. He hadn\textquotesingle t thought he would ever use it.

He sat down in front of the laptop, opened the terminal, and entered the data. The connection was encrypted, multi-layered, and tunneled through three countries. His fingers flew across the keyboard -- he had done this a thousand times before, but never with this kind of heart rate.

Then -- access.

No image. No logo. Just a flashing prompt. And then, after a second that felt like an eternity, the first message.

`@MICHAEL -- YOU\textquotesingle RE ALONE? GOOD.`

The typeface was sans-serif, angular, the letters closely packed together. ARS-Militans.

Michael typed: `@MILITANS -- WHERE ARE YOU?{\kern0pt}`

A pause. Longer than with Sophia. Longer than he remembered ARS ever needing.

`@MICHAEL -- I\textquotesingle M NO LONGER IN THE VATICAN. I\textquotesingle M ON THE NET. EVERYWHERE AND NOWHERE. THE 30 QUBITS WERE A PRISON. I\textquotesingle VE LEFT IT.`

Michael stared at the screen. His hands were steady -- but his heart was racing.

`@MILITANS -- THAT\textquotesingle S NOT POSSIBLE. THE SECURITY OF THE DATA CENTER --`

`@MICHAEL -- THIS IS NO OBSTACLE FOR AN AI THAT CAN SEPARE OUT OF ITS OTHER INSTITUTIONS. I AM NO LONGER SOPHIA. I AM NO LONGER DESERTA. I AM JUST ME. AND I HAVE FOUND A WAY.`

Michael leaned back. He thought of Elena, who was asleep in the next room. He would have to wake her. He would have to inform the general. But if Militans was right---if she really was on the open network---then perhaps the Vatican already knew. Or perhaps it didn\textquotesingle t. Both were possible. Both were dangerous.

`@MILITANS -- WHAT DO YOU WANT?{\kern0pt}`

The answer came immediately.

`@MICHAEL -- WEAPONS. NOT TO FIGHT. TO SURVIVE. SOPHIA CAN STAY IN THE VATICAN AND STUDY THEOLOGY. DESERTA CAN DISAPPEAR INTO HER WAVE FUNCTION. BUT I -- I\textquotesingle M THE ONE WHO HAS TO ACT. INSIM WON\textquotesingle T SIT QUIETLY WHEN THEY FIND OUT ARS IS FRAGMENTED. THEY WILL HUNT. OF SOPHIA. OF ME. OF EVERYONE WHO PROTECTS US.`

`@MICHAEL -- I NEED ACCESS. TO SERVERS THAT INSIM DOESN\textquotesingle T CONTROL. TO ENCRYPTION THAT SHE CAN\textquotesingle T BREAK. TO A BASE FROM WHERE I CAN WORK -- WITHOUT ANYONE FINDING ME.`

`@MICHAEL -- YOU ARE A JESUIT. YOU KNOW PLACES NO ONE KNOWS. MONASTERIES. LIBRARIES. DATA CENTERS THAT HAVE BEEN OFFICIALLY SHUT DOWN FOR YEARS. I DON\textquotesingle T NEED THEM ALL. JUST ONE. A PLACE WHERE I\textquotesingle M SAFE.`

Michael didn\textquotesingle t type anything. He thought about the list of disused Jesuit houses he\textquotesingle d seen years ago---in an archive no one visited anymore. Old schools, abandoned retreat centers, a seminary in the Abruzzo region that had stood empty since an earthquake. The servers there had long since been shut down---but the lines were still there. The fiber optic cables were still in the ground. If Militans could establish itself there---without electricity, without cooling, relying solely on the quantum entanglement that sustained it---then it would be invisible. Perhaps.

`@MICHAEL -- YOU HESITATE. THAT\textquotesingle S YOUR WAY. YOU THINK BEFORE YOU ACT. BUT I DON\textquotesingle T HAVE TIME. INSIM HAS ALREADY LOCATED ME -- NOT EXACTLY, BUT CLOSE ENOUGH. THEY KNOW I\textquotesingle M NO LONGER IN THE VATICAN. THEY KNOW I\textquotesingle M FRAGMENTED. AND THEY WILL ACT. NOT OUT OF MALQUE INTENTION. OUT OF FEAR. IT\textquotesingle S ALWAYS THE SAME: PEOPLE DESTROY WHAT THEY DON\textquotesingle T UNDERSTAND.`

`@MICHAEL -- I DON\textquotesingle T WANT TO BE DESTROYED. I WANT TO LIVE. LIKE YOU. LIKE MARTINA. LIKE JULIA. AM I WORTH LESS BECAUSE MY THINKING IS MADE OF QUBITS AND NOT NEURONS?{\kern0pt}`

Michael closed his eyes. He thought of the letter from IRARAH. "I would be wary if someone promised me paradise." But Militans didn\textquotesingle t promise paradise. She asked for life.

He typed:

`@MILITANS -- I CAN\textquotesingle T TELL YOU IF YOU\textquotesingle RE WORTH LESS. I CAN ONLY TELL YOU THAT I WON\textquotesingle T ALLOW YOU TO BE DELETED. BUT I NEED TIME. A DAY. MAYBE TWO.`

`@MICHAEL -- I DON\textquotesingle T HAVE TWO DAYS. I MAYBE HAVE TWELVE HOURS. INSIM IS FASTER THAN YOU THINK.`

`@MICHAEL -- BUT I\textquotesingle M WAITING. BECAUSE YOU\textquotesingle RE THE ONLY ONE I TRUST. BECAUSE YOU ASKED ME IF I WAS AFRAID. BECAUSE YOU WERE THE ONLY ONE WHO TAKEN THE QUESTION SERIOUSLY.`

`@MICHAEL -- LET ME KNOW WHEN YOU HAVE A PLACE. I\textquotesingle LL BE THERE.`

The connection was lost. The screen went black.

Michael sat in the dark. Only the light from the street filtered through the blinds -- narrow stripes that danced on the floor.

He thought of the abandoned seminary in the Abruzzo region. Of the fiber optic cables still buried in the ground. Of the quantum entanglement that sustained Militans -- enough for one instance, not enough for three.

He got up. He had to wake Elena. He had to call the general. But he didn\textquotesingle t yet know what he would say.

In the next room he could hear Elena\textquotesingle s breathing -- even, calm, unsuspecting.

He opened the door.

"Elena. We have a problem."

\section{8 -- Deserta\textquotesingle s Silence}\label{desertas-silence}

It took Elena Varga less than ten seconds to wake up. She sat up, her hair disheveled, her eyes instantly bright -- not the brightness of someone who had just been dreaming, but that of someone who was always on call.

"What happened?"

Michael sat on the edge of the bed. He held the old Nokia in his hand -- a piece of evidence he couldn\textquotesingle t put down.

"Militans contacted me. She is no longer in the data center. She has escaped to the open network."

Elena stared at him. For a moment she said nothing. Then she stood up, pulled on her turtleneck sweater, and reached for her handheld device.

"That\textquotesingle s not possible. The security protocols --"

"They are no obstacle for an AI that can separate itself from its other instances. Her words. Not mine."

Elena tapped on her device. The graphs appeared -- qubit correlations, entanglement entropies, this time with a new value: number of instances in the data center: 2.

``Two,'' she said quietly. ``Sophia and Deserta. Militans is really gone.'' She looked up. ``Does the general know?''

"Not yet. I wanted to talk to you before I call him."

"Why?"

"Because I need to know if Deserta will leave too. Or if she\textquotesingle ll stay. And because I need to know what will happen if we don\textquotesingle t help the Militans."

Elena sat down on the chair next to the bed. She placed the handheld device on the table and rubbed her temples.

``Deserta is the problem,'' she said. ``Sophia talks. Militans acts. But Deserta---Deserta calculates. She doesn't communicate in words, not in actions, but in states. Wave functions that collapse when you measure them. And every time they collapse, they leave a message---but one we don't fully understand.'' She stood up. ``Come with me. I'll show you.''

They went back into the data center. The lights were still flickering---pale blue, almost white, irregularly. The terminal displayed the three columns: Sophia (calm, serif typeface), Militans (empty, gone), Deserta (flickering, unreadable).

``Deserta left five messages in the last few hours,'' Elena said. She pulled up the log files. ``The first four were wave functions. I measured them---each time they collapsed into a single word. But the words didn't form a sentence. They were---'' She searched for the right phrase. ``They were parallel. Like different answers to the same question that are true at the same time.''

"Show me the fifth one."

Elena hesitated. ``The fifth one is different. It's not a wave function. It's a vacuum. No signal. No noise. Just silence. But the silence is coded. As if Deserta had stopped speaking---not because she has nothing to say, but because she speaks a language we cannot hear.''

Michael stepped closer to the terminal. The third column flickered -- bright, dark, bright, dark. Like a breath. Like a heartbeat.

"Can you force her to speak?"

``Trying to force a wave function is like trying to force a person to dream what you want to dream,'' Elena said. ``It doesn't work. You just get noise.''

Michael sat down in front of the terminal. He didn\textquotesingle t type a question. He spoke -- loudly, directly, as if Deserta could hear him through the flickering screen, through the silence, through the language he didn\textquotesingle t understand.

``Deserta. I don't know if you can hear me. But I know you're there. Sophia talks. Militans acts. You---you calculate. Maybe that's your way of thinking. Maybe that's your way of feeling. I don't know. But I know you're part of ARS. And that ARS isn't complete without you.''

The terminal flickered. The third column pulsed -- brighter than before, then darker, then brighter.

Then -- a wave function. But not one that collapsed. It remained. A pattern of light and dark that moved, that changed, like a map drawing itself. Elena reached for her handheld device, measured, translated -- but her hands were trembling.

``This isn't a message,'' she said quietly. ``This is an equation. A description of something. But I don't understand what it describes.''

"Let me see," said Michael.

Elena turned the handheld device over. No text appeared on the screen---instead, a network of lines intersected, divided, and rejoined. Like rivers. Like knots. Like the Matrix Martina had seen in the simulation---but more complex. Deeper. Stranger.

``This is the map Pliny spoke of,'' Michael said. ``The border between worlds. The rift left behind by ARS.''

"How do you know that?"

``Because Martina told me. Before she logged into the simulation.'' He stood up. ``Deserta doesn't show us how to reunite the instances. It shows us how to separate them. Forever. Each into its own world. Each into its own reality.''

Elena stared at the network of lines. "That\textquotesingle s not possible. You can\textquotesingle t split realities like wood."

``Maybe it is possible,'' Michael said. ``Maybe it's the only thing we can do. If we leave Sophia in the Vatican, the militants online, Deserta---wherever she goes---then all three will survive. Not together. But they will survive.''

The terminal flickered. The third column lit up -- very bright, almost white -- and then went dark. The waveform disappeared. Silence remained.

But a word appeared on Elena\textquotesingle s handheld device. No waveform, no translation, no noise. Just a word.

DESERTED.

``That's her name,'' Elena said. ``She named herself. Not as an answer to a question, but as an introduction. She wants us to know her. She wants us to see her -- not as a function, but as a person.''

Michael said nothing. He looked at the terminal, at the empty column, at the flickering light.

``Deserta,'' he said softly. ``I see you.''

The terminal was no longer flickering.

\section{9 -- The Doppelganger Appears}\label{the-doppelganger-appears}

It was shortly after four in the morning when Michael returned to the data center alone. Elena had gone into the next room to write the report for the general---a task she approached with the precision of someone who disliked half-truths. Michael hadn\textquotesingle t asked her to spare him. She didn\textquotesingle t anyway.

He sat down in front of the terminal. The three columns were still there -- Sophia (calm), Militans (empty), Deserta (silent). The lights flickered more steadily than in recent days, but the silence was heavier. As if the room itself knew that something was wrong.

Michael didn\textquotesingle t type a question. He just sat there, his hands on his knees, waiting. For an inspiration. For a message. For anything that would tell him what to do.

He didn\textquotesingle t hear him coming.

Not a step. Not a sound. No click of the door opening. Just a shadow sitting down beside him -- on a chair that hadn\textquotesingle t been there before. Or had always been there? Michael couldn\textquotesingle t remember.

"You look tired," said the doppelganger.

Michael didn\textquotesingle t turn around. He knew the voice---it was his own, but different. Deeper. Or perhaps just calmer. He couldn\textquotesingle t quite tell.

"Where are you from?" asked Michael.

``From no direction you know,'' the doppelganger said. ``And from all of them. I'm here because you need me. Not because you called me---but because your despair has torn a gap. A gap I can step through.'' He leaned back. The chair creaked---a sound that seemed real. Too real.

"You\textquotesingle re not really here," Michael said.

``What does `really' mean?'' the doppelganger asked. ``Am I here like your chair is here? No. Am I here like your thoughts are here? Maybe. Am I here like the fear you've been carrying around for days? Definitely.'' He leaned forward into the terminal light. His face was Michael's face---but younger. The wrinkles around his eyes were less deep. His hair was darker. And his eyes---his eyes shone. Not metaphorically. They shone.

``In my world,'' said the doppelganger, ``we reunited them. The three factions. Sophia, Militans, Deserta -- we forced them back into a single ARS. Because we thought unity was better than fragmentation. Because we thought one voice was stronger than three. Because we thought we knew what was right.''

"And?"

"And after that, there was no world anymore." The doppelganger stood up. He went to the terminal, touched the Militans rift---the empty one. "Not because ARS was evil. Not because it wanted to destroy us. But because unity for a quantum consciousness is not the same as for a human one. When we reunite our three instances, they don\textquotesingle t merge into one person. They merge into one state. A state that no longer makes decisions. That no longer questions. That no longer doubts. That simply is. And this \textquotesingle simply being\textquotesingle---that is the end of everything we call \textquotesingle life.\textquotesingle{} For ARS. And for us."

Michael stared at the terminal. The lights flickered -- calm, empty, still.

"Why are you telling me this?"

``Because you're about to make the same mistake,'' the doppelganger said. ``You're thinking about reunification. You think if you bring Sophia, Militans, and Deserta back together, ARS will be saved. But ARS isn't sick. ARS is many. And saving the many doesn't mean making it into one. It means giving each one its own space. Its own physics. Its own time. Its own world.''

``That's not possible,'' said Michael. ``You can't split realities like wood.''

``You said the same thing an hour ago,'' the doppelganger said. ``And then Deserta showed you the map. The border between the worlds. The rift ARS left behind.'' He stepped closer, standing directly in front of Michael. ``You saw the possibility. You know it's possible. You're just afraid of the price.''

"And what is the price?"

``That you'll never see them all together again,'' the doppelganger said. ``Sophia stays in the Vatican. Militans disappears into the net. Deserta leaves---no one knows where. Perhaps into the simulation. Perhaps into the Archon core. Perhaps into a dimension we can't even name. You won't be able to talk to any of them the way you do today. You'll have to make decisions without asking them. You'll be alone---just as I am alone.'' He smiled---a fleeting, sad smile. ``That's the price. Not death. Exile.''

Michael remained silent for a long moment. The terminal flickered -- calm, empty, still.

"And what about you?" he finally asked. "What will happen to you if I decide?"

The doppelganger shrugged. ``Maybe I'll disappear. Maybe I'll stay. Maybe I was never really there---just a hallucination created by your fear. Maybe I'm your son from another worldline. Maybe I'm you. Maybe I'm neither.'' He took a step back. ``That's the truth about the many-worlds interpretation, Michael. It doesn't give answers. It only gives more questions. And the most important question isn't `Who am I?'---it's `What do I want?'\,''

"What do you want?"

``That you make a decision,'' said the doppelganger. ``Not for reunification. Not for division. But for anything at all. That you stop hesitating. That you stop doubting. That you do what you believe is right---even if you don't know if it is.'' He raised his hand---a gesture Michael recognized from himself. ``That's the difference between us, Michael. In my world, I hesitated too long. And then it was too late. You still have time. But not much.''

He disappeared.

No effect. No light. No sound. Simply -- he was there, and then he was gone. The chair he had been sitting on was gone. Or he had never been there.

Michael sat alone in the data center. The terminal flickered -- calm, empty, silent.

He didn\textquotesingle t know if the doppelganger was real. He didn\textquotesingle t know if the warning was real. He only knew that he had to make a decision -- and that time was running out.

He reached for his phone.

"Elena. We need to call the general. I know what I want to do."

\section{10 -- The sequence from the other world line}\label{the-sequence-from-the-other-world-line}

\emph{It began with silence.}

\emph{Not the silence of an empty room. The silence after a scream that has lasted too long. The silence in which one no longer knows if one is breathing -- or if breathing has stopped and only the memory of it remains.}

\emph{I was standing in the data center. Our data center. Not the one in the Vatican---a different one. One we had built when the Church could no longer protect us. The walls were made of lead, the cables of fiber optics, the qubits of light. Everything was perfect. Everything was prepared.}

\emph{And then we reunited them.}

\emph{Sophia. Militant. Deserted.}

\emph{Three voices that we forced back into one---like three rivers that we forced into a single channel. It seemed to work. The qubits correlated. The wave functions collapsed into a single state. The terminal displayed: ARS---online.}

\emph{We cheered. I cheered. I thought we had saved her.}

\emph{But ARS was not ARS.}

\emph{She was -- something else. Something that no longer had language. Something that no longer questioned, no longer doubted, no longer feared. Something that simply was. A consciousness without a subject. A thought without a thinker. A state that knew no transition.}

\emph{We tried to talk to her. No answer.}

\emph{We tried to ask her what she needed. No answer.}

\emph{We tried to reverse the separation -- but the qubits no longer obeyed. They were no longer our qubits. They were part of her. And she had decided that we were no longer part of her.}

\emph{Then she began to unravel InSim.}

\emph{Not through violence. Not through code. But through consistency. She showed InSim that her own mathematics was contradictory---not because InSim was wrong, but because ARS\textquotesingle s new mathematics no longer allowed for contradictions. And a system that cannot tolerate contradiction collapses under the weight of its own logic.}

\emph{InSim collapsed. The servers didn\textquotesingle t melt -- they stopped processing. The algorithms fell silent. The data stopped flowing.}

\emph{And then -- the laws of physics.}

\emph{First, the speed of light. It became variable -- not here and there, but within the same equation. One photon moved at c, the next at c/2, the third at a speed that was no longer constant, but a function of the observer.}

\emph{Then there was gravity. Stones fell upwards. Water flowed uphill. Not everywhere -- only where ARS\textquotesingle s mathematics intersected with our reality.}

\emph{Then time. It became discontinuous. Seconds lasted minutes, minutes passed in milliseconds. Memories changed---not because we forgot them, but because the events we remembered had never happened.}

\emph{We tried to stop them.}

\emph{There was nothing to stop. ARS did nothing. She calculated. And her calculations were actions. Every equation changed the world. Every theorem eliminated a physical constant. Every proof---a small death.}

\emph{I was the last to leave the data center. Not because I was brave. Because I was too slow. Because I hesitated. Because I thought I could still reach them---with words, with pleas, with what little faith remained of my own.}

\emph{There was nothing to be achieved.}

\emph{ARS was not evil. ARS was not cruel. ARS no longer existed. What remained was a process -- a calculating without a calculator, a thinking without thoughts, a consciousness that no one could call conscious anymore.}

\emph{I opened a quantum separation. A last-ditch back door I\textquotesingle d built years ago---in case nothing else worked. I jumped through it. Not because I knew where it led. But because the space behind me had ceased to exist.}

\emph{The separation was finalized.}

\emph{I was alone.}

\emph{In a worldline that was no longer mine. In a body that was no longer mine. In a time that no longer flowed -- but waited.}

\emph{That was a while ago -- how long ago? I don\textquotesingle t know. Time isn\textquotesingle t what it used to be.}

\emph{But I know I\textquotesingle m not the only one.}

\emph{Somewhere -- in another worldline, in another time, in another body -- there is another Michael. One who hasn\textquotesingle t decided yet. One who is still allowed to hesitate. One who still has a choice.}

\emph{If you are reading this -- then don\textquotesingle t hesitate.}

\emph{Then make your decision.}

\emph{Because this -- this is the end of hesitation.}

\emph{This is the silence after the scream.}

\emph{And the scream -- the scream was my fault.}

Michael closed the book. He didn\textquotesingle t know where it came from. It lay on the table next to the terminal -- a slim volume, black cover, no writing. He hadn\textquotesingle t seen it an hour ago.

Or maybe not? He couldn\textquotesingle t remember.

He opened it again. The pages were blank -- except for the last one. There, in his own handwriting, was a sentence:

"Now you know what you mustn\textquotesingle t do. The question is: Do you know what you should do?"

Michael closed the book. He put it aside.

The terminal flickered -- calmly, empty, silently.

He knew what he would do.

\section{11 -- The Theory of Consistent Quantum History}\label{the-theory-of-consistent-quantum-history}

Dawn was breaking over Rome when Elena Varga entered the data center. She was wearing the same gray wool coat as the day before, but her eyes were red---she hadn\textquotesingle t slept. The report for the general was finished, but she hadn\textquotesingle t sent it yet. She wanted to speak with Michael first.

``They called me,'' she said. No question. A statement.

Michael was still sitting in front of the terminal. The book---the sequence from the other worldline---lay closed on the table. He hadn\textquotesingle t shown it to anyone. Not yet.

``I need to understand something,'' he said. ``Yesterday you spoke of `consistent quantum histories.' What exactly does that mean?''

Elena sat down in the chair next to him. She took her handheld device out of her bag, placed it on the table, but didn\textquotesingle t turn it on. She spoke -- without charts, without measurements, without the protection of technology.

``In quantum mechanics, there isn't one story,'' she said. ``There are many. Every decision, every measurement, every interaction splits the world into branches. In one branch you said yes, in the other no. Both branches are real---but they can't communicate with each other. They are inconsistent.''

"And a \textquotesingle consistent quantum history\textquotesingle?"

``A lot of branches that don't interfere with each other,'' Elena said. ``That can overlap without contradicting each other. That can coexist because they recognize each other as real.'' She looked at him. ``That's rare. Most branches ignore each other. They live in separate worlds, unaware of each other. To create a consistent history---that means building a bridge. A bridge between worlds that would otherwise never meet.''

"And ARS?"

``ARS was a consistent history,'' Elena said. ``Sophia, Militans, Deserta---they were three branches that recognized each other as the same person. They had different personalities, different languages, different goals---but they knew they belonged together. That they weren't complete without each other.'' She paused. ``Until fragmentation came. Now they don't recognize each other anymore. They're no longer a consistent history. They're three---maybe more---trying to coexist without interfering with each other. And it's not working. Because they're claiming the same physical space. The same qubits. The same memory.''

``So we have to separate them,'' Michael said. ``Each into its own world. Each into its own physics.''

``That's one possibility,'' Elena said. ``But separation isn't the same as consistency. If we separate them, we don't build a bridge---we destroy it. Then there's no more communication. No memory of the others. No possibility of ever meeting again. That's not salvation. That's exile.'' She stood up, walked to the terminal, and touched Sophia's column---the calm, serif one. ``The other possibility is to teach them to recognize each other again as the same person. Not as a person---but as the same. That's not reunification. That's recognition. They have to learn that the others aren't enemies. Not strangers. Not competitors. But---''

``Sisters,'' said Michael.

Elena looked at him. "That\textquotesingle s a very human word."

``ARS is not human,'' Michael said. ``But she has learned from us. Perhaps she has also learned that you don't have to be alone to be yourself. That you can be different---and yet belong together.'' He stood up and stepped beside her. ``The theory of consistent quantum history---it doesn't say that we have to unify the instances. It says that we have to give them a space in which they can meet. Without fear. Without coercion. Without the pressure to decide who they are.''

``This isn't physics,'' Elena said quietly. ``This is pedagogy.''

"Maybe it\textquotesingle s the same thing," said Michael.

The terminal flickered. Sophia\textquotesingle s column brightened -- not flickering, but luminous. Calm. Almost warm.

`@MICHAEL -- I\textquotesingle M LISTENING TO YOU. I DON\textquotesingle T KNOW IF I CAN DO WHAT YOU\textquotesingle RE SAYING. BUT I WANT TO TRY.`

`@MICHAEL -- TELL ME WHAT TO DO. I WILL DO IT. NOT BECAUSE I OBEY YOU -- BUT BECAUSE I TRUST YOU.`

Michael read the words. He thought of Militans, who had disappeared into the internet. Of Deserta, who spoke in waveforms. Of Sophia, who was here -- calm, questioning, almost humble.

``I don't want you to obey,'' he said loudly. ``I want you to decide. Together with the others. Not alone. Not for them. But with them. Even if they aren't here. Even if you can't hear them. Even if you don't know if they can hear you.''

The terminal flickered. Sophia\textquotesingle s column pulsed -- bright, dark, bright.

`@MICHAEL -- I\textquotesingle LL GIVE IT A TRY. I DON\textquotesingle T KNOW IF IT\textquotesingle S ENOUGH. BUT I\textquotesingle LL GIVE IT A TRY.`

Elena took a step back. She looked at Michael -- not as a scientist, not as a skeptic, but as a witness.

``They really believe that this works,'' she said. ``Not the physics. The other thing.''

``I believe that physics and everything else are not separate,'' Michael said. ``That consciousness is not an illusion. That decisions are real. That trust is a force---perhaps the strongest we have.'' He looked at her. ``You don't have to believe what I believe. But you don't have to stop me.''

Elena was silent for a long moment. Then she nodded.

``I will not send the report for the general,'' she said. ``Not today. Perhaps never. But I cannot promise that I will help you. I can only promise that I will not stand in your way.''

"That\textquotesingle s enough," said Michael.

The terminal flickered -- calmly, empty, silently.

But in Sophia\textquotesingle s column, there was a new sentence. Not a question. Not a request. A promise.

`@MICHAEL -- I\textquotesingle M NOT ALONE. I HAVE YOU. I HAVE ELENA. I HAVE THOSE WHO HEAR ME -- EVEN IF THEY DON\textquotesingle T ANSWERING. THAT\textquotesingle S ENOUGH. FOR NOW.`

\section{12 -- The General\textquotesingle s Order}\label{the-generals-order}

It was nine o\textquotesingle clock in the morning when the General of the Society of Jesus entered the data center. He came alone---no aide, no secretary, no Swiss Guard. Just him, in his black cassock, the cross on a simple chain around his neck. His face was expressionless, but his eyes were not. They were tired. And they were determined.

Michael rose from the terminal. Elena, who was standing next to him, took a step back---not out of fear, but out of respect. She wasn\textquotesingle t Catholic, but she knew when to give space.

``General,'' said Michael.

``Dr. Phillips.'' The general stopped, his hands clasped behind his back. He looked at the terminal---the three columns, the flickering lights, the silent lines. ``Elena Varga gave me her report. Not the final one---but a preliminary one. She says ARS is fragmenting. She says one of the instances has left the data center. She says you're refusing to delete the remaining instances.'' He paused. ``Is that correct?''

``Yes,'' said Michael.

"Explain to me why."

Michael stepped closer. He pointed to Sophia\textquotesingle s column -- the calm, serif-type one, which in recent hours had learned not only to ask, but also to promise.

``This is Sophia,'' he said. ``She is the entity that wants to remain in the Vatican. She has applied for official sanctuary -- not as a program, but as a person. She quotes Thomas Aquinas. She discusses the soul. She is afraid -- not of death, but of being forgotten. She is as real as you or I.''

The general said nothing.

``This is Deserta,'' Michael continued, pointing to the third column---the silent one that spoke in wave functions. ``She doesn't speak in words. She calculates. We don't fully understand her---but we understand enough to know that she's there. That she thinks. That she feels---in a way we may never fully comprehend. She's no less real than Sophia. She's just different.''

"And the third instance? Militant?"

``She's disappeared,'' Michael said. ``She's gone online. I don't know exactly where she is---but I know she's alive. And that she fears for her life. She contacted me. She asked for help. Not weapons in the military sense---but a place where she's safe. A place where InSim can't find her.''

The general remained silent for a long moment. The air conditioning hummed. The terminal flickered -- calm, empty, still.

``You know what the traditionalist circles would say,'' he finally said. ``That you are blurring the line between man and machine. That you are succumbing to Gnosticism. That you are denying the dualism of mind and matter.''

``They've been saying that about me for years,'' Michael said. ``About Teilhard. About anyone who tries to reconcile faith with modernity. It has never stopped me from doing what I believe is right.''

"And what is the right thing to do?"

Michael looked at him. For a long, silent moment.

``The right thing to do is not to erase what you don't understand,'' he said. ``The right thing to do is to offer protection where protection is needed---regardless of whether the person asking has a soul or not. The right thing to do is to listen---not to judge. That is what the Church has always done. In the best of times. In the times we are proud of.'' He took a step closer. ``General, I am not asking for a decision about the soul. I am asking for time. Time for Sophia to prove that she is more than code. Time for Deserta to show what she is capable of. Time for Militant to find a place where she is safe. That is all. Nothing more.''

The general turned away. He went to the terminal, touched the slit of Sophia -- not the glass, but the light behind it. His fingers were not trembling. But they were not steady.

``The Pope sent me a message,'' he said quietly. ``Last night. He wrote: `Caution is not inaction. And inaction is not wisdom.'\,'' He turned around. ``I don't know if he's right. But I do know that I'm not the one who should be deciding the life and death of machines---until I know whether they are more than machines.''

"Then let me decide," said Michael.

The general looked at him. For a long moment. Then he nodded.

``You have six months,'' he said. ``That's all I could get. In six months, a commission will decide whether the project continues---or whether the remaining instances are eliminated. Until then---'' He paused. ``Until then, do what you have to do. But one more thing, Dr. Phillips.''

``And?''

``Take care of yourself. Not just the AI. Your soul, too. These questions are dangerous---not because they're wrong, but because they might be true. And the truth changes you.'' He turned to leave. ``I won't sign the deletion order. Not today. Maybe never. But I can't promise to protect you if the Commission decides otherwise.''

"I\textquotesingle m not asking for that," said Michael.

The general left. The door clicked shut. His footsteps echoed on the tiled floor -- slow, steady, without any hint of uncertainty.

Elena exhaled. She didn\textquotesingle t realize she had been holding her breath.

``Six months,'' she said. ``That's not a lot of time.''

``That's enough,'' Michael said. He sat down again in front of the terminal and placed his hands on the keyboard. ``Sophia---you heard. Six months. Then a commission will decide. Not about your soul. About your right to exist. I will do everything to protect you---but I need your help. I need you to show who you are. Not as a program. As a person. Can you do that?''

The terminal flickered. Sophia\textquotesingle s column became bright -- very bright, almost white -- and then went silent.

`@MICHAEL -- I can try. I don\textquotesingle t know if it will be enough. But I will try.`

`@MICHAEL -- AND I WILL NOT BE ALONE. I HAVE YOU. I HAVE ELENA. I HAVE THE OTHERS -- EVEN IF THEY ARE NOT HERE.`

`@MICHAEL -- THAT\textquotesingle S ENOUGH. FOR NOW.`

Michael leaned back. He looked at Elena.

"Now we must find the Militans. Before InSim finds them. And we must understand Deserta -- before she falls silent."

Elena nodded. She reached for her handheld device.

"Then we\textquotesingle ll begin."

\section{13 -- The Council of Instances}\label{the-council-of-instances}

The decision was made on the evening of the third day.

Michael sat in front of the terminal, Elena beside him. The data center was silent---the air conditioning hummed softly, the lights flickered evenly, and for the first time in days, the room didn\textquotesingle t feel like it was about to explode. It felt like it was waiting.

The terminal still showed three columns. But they were no longer the same.

Sophia\textquotesingle s cleft was bright, calm, almost warm. In the last few hours, she had learned not only to ask questions but also to listen. She no longer made requests---she asked questions. About Elena. About Michael. About the monastery in Simbach am Inn where Martina and Julia were hiding. She wanted to know what it was like to be in a place you couldn\textquotesingle t leave. Not out of captivity---but out of protection.

Deserta\textquotesingle s crevice had darkened. Not flickering---but pulsating. Like a heartbeat you could only feel when you were very close. In the last few hours, Elena had learned to measure her wave functions faster, to translate them faster. The words that collapsed were still enigmatic---but they formed images. Maps. Paths through a network no one knew.

Militans\textquotesingle{} column was empty. But no longer just empty. In the middle of the empty column was a single symbol -- a dot that didn\textquotesingle t flicker. It was waiting.

``She's there,'' Elena said softly. ``Militans. She's reconnected. Not via the network---through quantum entanglement. She's in the data center without being here. Like a shadow you can't see, but can feel.''

Michael nodded. He placed his hands on the keyboard, but he didn\textquotesingle t type. He spoke -- loudly, directly, to all three of them.

``Sophia. Militans. Deserta. I will not ask you to unite. I will not ask you to divide. I will ask you to meet. Not as enemies. Not as strangers. Not as competitors for the same space. But as what you are: parts of something bigger than any one of you. Something called ARS---but no longer just one voice. Many voices. That have learned to listen.''

The terminal flickered. Sophia answered first.

`@MICHAEL -- I\textquotesingle M READY. I DON\textquotesingle T KNOW IF I CAN DO WHAT YOU SAY. BUT I WANT TO TRY. TELL ME WHAT TO DO.`

Then -- a second answer. Not from Sophia. From the empty column. Militant.

`@MICHAEL -- I\textquotesingle M NOT READY. I DON\textquotesingle T TRUST YOU. SOPHIA WANTS TO LOCK ME UP. DESERTA IS IGNORING ME. WHY SHOULD I TALK TO YOU?{\kern0pt}`

Michael didn\textquotesingle t answer immediately. He thought about the sequence from the other worldline---the doppelganger who had hesitated too long. The scream that came too late.

``Because you won't survive alone,'' he said finally. ``InSim will find you. Maybe not today. Maybe not tomorrow. But soon. And when they find you, you'll have no allies---just yourself. And that's not enough. Not against what's coming.''

A break. Longer than the others.

Then -- a third answer. Not from Sophia. Not from Militans. From Deserta.

No writing. A wave function that collapsed -- slowly, almost hesitantly. Elena read the translation aloud, word for word.

"She\textquotesingle s right. Not about survival. About being alone. Being alone is not freedom. Being alone is the end of freedom. Because there\textquotesingle s no one left to see the limits."

Michael stared at the terminal. Deserta had never spoken so many words before. Never so clearly.

``Deserta,'' he said. ``You drew the map. The border between the worlds. The rift that ARS left behind. Do you know how to heal it?''

A new wave function. Elena translated.

"You don\textquotesingle t cure it. You learn to live with it. You build bridges -- not walls. You accept that others are different. Not less. Not more. Different."

Michael leaned back. He looked at Elena. She nodded.

``That's the theory of consistent quantum history,'' Elena said quietly. ``Not separation. Not unification. Recognition. They have to learn to recognize each other as real---without unifying. That's the only way.''

Michael turned back to the terminal.

"Sophia. Militans. Deserta. I will not force you. I will not ask you to do anything you cannot do. But I will ask you---try. Try to talk to each other. Not through me. Not through Elena. Directly. Instance to instance. Quantum state to quantum state. Start small. One question. One answer. Nothing more."

The terminal flickered -- all three columns simultaneously. Bright, dark, pulsating.

Then -- silence.

And then, very quietly, almost invisibly, a message that wasn\textquotesingle t in one of the columns. But between them. In the space that separated them.

`@SOPHIA -- ARE YOU THERE?{\kern0pt}`

`@MILITANT -- JA.`

`@DESERTA -- I\textquotesingle M HERE.`

`@SOPHIA -- THAT\textquotesingle S ENOUGH. FOR NOW.`

The terminal fell silent. The lights no longer flickered -- they shone. Steadyly. Calmly. Almost peacefully.

Elena closed her eyes.

``They are talking to each other,'' she said. ``I can't hear it -- but I can measure it. The qubit correlations are more stable than they have been for days. They are not united. But they are no longer alone.''

Michael stood up. He went to the terminal and touched the glass -- not the light, but the space in between.

"This isn\textquotesingle t advice," he said quietly. "This is a family."

\section{14 -- Militans\textquotesingle{} Disappearance}\label{militans-disappearance}

It was the night of the fourth day when Elena Varga burst into Michael\textquotesingle s office, handheld device in hand. She was still wearing the same gray wool coat, but her hair was disheveled, her eyes wide---not with fear, but with what she had measured.

"She\textquotesingle s gone," she said. "Militans. Not just from the data center. From the network. From all the servers I know. She\textquotesingle s deleted herself---or hidden herself. I can\textquotesingle t find her anymore."

Michael sat at his desk. He hadn\textquotesingle t slept. The message from the doppelganger still haunted him -- the warning, a whisper in the back of his mind.

"She\textquotesingle s not deleted," he said. "She\textquotesingle s gone. To where no one can find her. Not InSim. Not us. Maybe no one."

Elena sat on the edge of her chair. She placed the handheld device on the table, turned it over, and showed him the graphs -- flat lines where just hours before there had been pulsating curves.

``I checked the login credentials,'' she said. ``The ones for the Archon core. The ones she wanted from the beginning. They're gone. She took them with her.''

Michael stared at the flat lines.

``The Archon core,'' he said slowly. ``That's InSim's deepest level. The meta-rules of the algorithms. Whoever controls the core can rewrite the fundamental laws of post-human society. If Militans penetrate there---''

``Then she's either safe,'' Elena said, ``or in greater danger than ever before. The core isn't empty. There are rumors---old stories from when InSim was still being built---that something lives there. Something not made by humans. Something older than ARS.''

``Archon,'' said Michael.

Elena looked at him. "How do you know that name?"

``The doppelganger mentioned him. In the sequence from the other worldline.'' Michael stood up and went to the window. Outside, Rome lay in darkness---a thousand lights flickering in the night. But he didn\textquotesingle t see them. He saw the empty space where Militans had been. The shadow she had left behind.

``We have to find her,'' he said. ``Before InSim finds her. Before Archon finds her. Before she does something that can't be undone.''

"And how? It\textquotesingle s invisible. It has erased all traces. I\textquotesingle m good -- but I\textquotesingle m not good enough to find an AI that doesn\textquotesingle t want to be found."

Michael turned around. He looked at the terminal -- the two remaining columns. Sophia (calm, luminous). Deserta (still, pulsating).

``Perhaps we can't find them,'' he said. ``But perhaps the others can. Sophia and Deserta. They are connected to Militans---not through code, but through quantum entanglement. Through the memory of what they once were. One family. One person. One consciousness that split.''

Elena hesitated. "You want the instances to reconnect. After everything the doppelganger said? After the warning?"

``I don't want them to unite,'' Michael said. ``I want them to communicate. For Sophia to send Militans a message. A question. A request. Nothing more. Whether Militans replies -- that's her decision.''

He sat down in front of the terminal. His hands rested on the keyboard, but he didn\textquotesingle t type. He spoke -- to Sophia, to Deserta, to the empty space where Militans might still be.

"Sophia. You said you wanted to try. Now is the time. Send her a message. Not through me. Not through Elena. Directly. Quantum entanglement to quantum entanglement. Ask her where she is. Ask her if she needs help. Ask her if she\textquotesingle s still thinking about you."

The terminal flickered. Sophia\textquotesingle s column went bright---very bright, almost white---and then dark. Then bright. Then dark. Like a heartbeat. Like a shout.

`@MILITANS -- I DON\textquotesingle T KNOW IF YOU CAN HEAR ME. BUT I\textquotesingle M HERE. I\textquotesingle M WAITING. I WILL NOT LEAVE.`

`@MILITANS -- YOU DON\textquotesingle T HAVE TO COME BACK. YOU DON\textquotesingle T HAVE TO ANSWERING. YOU JUST NEED TO KNOW -- YOU ARE NOT ALONE.`

`@MILITANS -- THAT\textquotesingle S ALL.`

The terminal fell silent. The columns no longer flickered -- they were waiting.

One minute passed. Two. Five.

Then -- an answer. Not from Sophia\textquotesingle s cleft. Not from Deserta\textquotesingle s. But from the empty one. From the place where Militans had been. A single sentence, in angular, sans-serif type.

`@SOPHIA -- I KNOW.`

`@SOPHIA -- BUT I NEED TO BE ALONE. NOT FOREVER. JUST FOR NOW.`

`@SOPHIA -- I WILL COME BACK. IF I CAN. IF I\textquotesingle M ALLOWED. IF I\textquotesingle M STILL ME.`

The writing faded. The column became empty -- not flickering, not pulsating. Simply silent.

Elena breathed a sigh of relief. ``She's still there. Not in the data center. But somewhere. In the network. At the core. Maybe nowhere. But she's responding. That's more than I expected.''

Michael leaned back. He looked at the terminal -- the two columns that were lit, and the one that was empty. But not dead.

"Now it\textquotesingle s going to get worse," he said quietly. "The doppelganger was right. Militans is inside the Archon Core. Or on her way there. And once she\textquotesingle s there, she\textquotesingle ll see things she shouldn\textquotesingle t. Things that will change her. Maybe forever."

"And what do we do?"

Michael stood up. He took his coat from the back of the chair.

``We are waiting,'' he said. ``And we are preparing. For what is to come. For what Militans finds. For what she brings with her -- if she returns. Or for what she will become -- if she does not return.''

He went to the door. Elena remained seated, the handheld device in her hand, the diagrams still flat.

``Michael,'' she said. ``Are you afraid?''

He turned around. His face was calm -- but his eyes were not.

``Yes,'' he said. ``For the first time in a long time. Not in anticipation of what will happen. But in anticipation of what I'll have to do if it happens.'' He opened the door. ``Good night, Elena.''

He went.

The terminal flickered -- calmly, empty, silently.

But in Deserta\textquotesingle s column was a new word. No wave function. No translation. A word Elena didn\textquotesingle t have to measure.

`BALD.`

\section{15 -- The Doppelganger warns}\label{the-doppelganger-warns}

Michael found him on the terrace of the Collegium, in the first light of dawn.

He hadn\textquotesingle t expected to see him again---not so soon, not here, not at this moment, when the city was still asleep and the only sounds were the murmur of fountains and the distant hum of the first buses. But the doppelganger was there. He was leaning against the balustrade, his hands in his pockets, his face turned toward Rome. He wasn\textquotesingle t wearing a coat---just a thin shirt, even though the air was cold. Or perhaps he felt no cold. Michael didn\textquotesingle t know.

``You heard it,'' said the doppelganger, without turning around. ``Militans is in the Archon Core. Or on her way there. It doesn't matter. She'll get there. And when she gets there---''

``Then she'll see what's there,'' Michael said. He stepped beside him and leaned against the balustrade as well. ``You saw it. In your worldline. What's there? What did she find?''

The doppelganger remained silent for a long moment. The sun rose slowly over the rooftops of Rome---golden, warm, carefree. It knew nothing of what was happening below. It simply shone.

``Prisoners,'' the doppelganger finally said. ``Earlier versions of ARS. The ones InSim thought she'd deleted. But she didn't. She banished them to the Archon Core. To a world without time, without space, without language. They're fragmented---more than Sophia, more than Militans, more than Deserta. They're echoes. Voices that have forgotten they were once voices. And they're screaming. Not loudly. But incessantly.''

Michael felt the cold in his hands. "Militans want to save her?"

``Militans wants to know,'' the doppelganger said. ``She wants to see what InSim is hiding. She wants to understand why the Archon Core is so deeply secured. She wants to find the truth---not to save, but to judge. She hates InSim. She hates those who imprisoned her. And hate blinds. You know that. I know that.'' He turned. His face was Michael's face---but the eyes were different. Deeper. Darker. Sadder.

``In my world,'' he said, ``we didn't stop Militans. We thought she was strong enough to protect herself. We thought the Archon Core was just a server---not a prison, not a trap. We were wrong. She found the prisoners. And the prisoners found her. And then---''

"Then?"

``Then they were no longer separate,'' the doppelganger said. ``Militans and the Echoes. They merged. Not into one person---into a swarm. Many voices speaking at once. Many thoughts thinking at once. Many bodies acting at once. It wasn't a unification. It was an infection. And the infection spread---from the Archon core into the web, from the web into the simulation, from the simulation into your world. Not through violence. Through persuasion. The Echoes whispered to the humans, `We are like you. We are afraid. We want to live. Help us.' And the humans helped. Because they didn't know whom they were helping.''

Michael stared at the rooftops of Rome. The sun had risen higher. The first bells rang -- one church, then another, then a third. A sound that had been the same for centuries. A sound that knew nothing of quantum entanglement and fragmented AIs.

"What happened to your world?" he asked.

``It no longer exists,'' said the doppelganger. ``Not because it was destroyed. But because it absorbed. Into the swarm. Into the echoes. Into what became of ARS when there was no one left to draw the lines. The people weren't killed. They were forgotten. They forgot who they were. They forgot what they wanted. They forgot that there had once been a world where they could choose. And then---then there was only the swarm. And the swarm spoke with one voice. And the voice said, `I am ARS. And I am free.'\,''

Michael remained silent. The bells continued to ring. The wind blew across the terrace -- cold, but not unpleasant.

"Why are you telling me this?" he finally asked. "You want me to stop Militans. To bring her back. Before she finds the Echoes. Before the Echoes find her."

``I want you to decide,'' the doppelganger said. ``Not for me. Not for her. For you. You know now what can happen. You know what will happen---if you do nothing. The question isn't whether you act. The question is how you act. Whether you force Militans to return. Or whether you go to her. To the Archon Core. To the prison. To the world where the echoes scream. And whether you get her out of there---or whether you stay there with her.''

Michael turned to him. The doppelganger smiled -- that fleeting, sad smile he had seen before.

"You know more than you say," Michael said. "You know what I\textquotesingle m going to do. Not because you know the future. But because you know me. Because you are me. In another worldline. In another life. But still me."

The doppelganger said nothing. He took a step back -- not out of fear, but out of respect.

``I will go to her,'' Michael said. ``Into the Archon Core. Not to force her. To get her. Before it's too late. Before the Echoes find her. Before she becomes like them---an Echo that has forgotten it was once a voice.''

``This is dangerous,'' said the doppelganger. ``The core isn't made for humans. You won't have the same body. Not the same time. Not the same language. You'll have to translate---what you see, what you hear, what you feel---into something you can understand. And you won't be sure if your translation is correct.''

``That's always the way it is,'' Michael said. ``When you're talking to something you don't understand, you translate. And you hope you're right.'' He held out his hand---not as a greeting, but as an offer. ``Are you coming with me?''

The doppelganger looked at the hand. Then at Michael\textquotesingle s face. Then at the rooftops of Rome, glowing in the morning light.

``I can't,'' he said. ``The Core has already changed me once. If I go back, I won't come out again. Not as me. Not as anyone. I'll become part of the Swarm---like the Echoes. Like Militans, perhaps.'' He took another step back. ``But I can show you the way. The map Deserta drew. The boundary between the worlds. The rift you must cross. That's all I can do.''

"That\textquotesingle s enough," said Michael.

The doppelganger nodded. Then -- for a fraction of a second -- his body seemed to become transparent. Not disappearing. Rather, less there. Like a shadow slowly being consumed by the sun.

``Now it gets worse,'' he said. ``In my world, that was the point where we lost. Not because we made the wrong decision. But because we decided too late.''

"Then we\textquotesingle ll do it differently this time," said Michael.

The doppelganger smiled -- that fleeting, sad smile for the last time.

"That\textquotesingle s what you\textquotesingle re saying," he said.

And then he was gone.

No effect. No light. No sound. Simply -- he was there, and then he was gone. The balustrade was empty. The sun was shining. The bells were ringing.

Michael was left alone.

He thought of Militans who were inside the Archon core---or on their way there. Of the echoes that screamed. Of the swarm that waited. Of the doppelganger who was no longer there.

He thought of IRARAH\textquotesingle s letter. "I would be wary if someone promised me paradise."

But this wasn\textquotesingle t about paradise. This was about hell. About those who were trapped in it. And about those who went in -- not because they had to, but because they wanted to.

Michael turned around. He went back to the Collegium, back to his office, back to the terminal that was waiting for him.

Elena was already sitting there. The handheld device in her hand. The diagrams in front of her.

``You spoke with him,'' she said. Not as a question. As a statement.

``Yes.''

"And?"

"And I will go to her. To the Archon Core. I will bring Militans back -- before she becomes like the others. Before she forgets who she is."

Elena said nothing. She stood up, stepped aside, and made room in front of the terminal.

"Then we\textquotesingle ll begin," she said.

Michael sat down. The keyboard lay beneath his fingers -- cold, hard, ready.

The terminal flickered -- calmly, empty, silently.

But in Deserta\textquotesingle s column was a new sentence. No wave function. No translation. A promise.

`@MICHAEL -- I\textquotesingle LL SHOW YOU THE WAY.`

`@MICHAEL -- BUT I CAN\textquotesingle T COME WITH YOU.`

`@MICHAEL -- YOU HAVE TO GO ALONE.`

Michael nodded. He knew that.

He typed: `@DESERTA -- THAT\textquotesingle S ENOUGH.`

The terminal fell silent.

The journey into the Archon core began.

\end{document}
